---
title: "Acquiring and cleaning zika data"
author: "Carson Sievert"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  fig.width = 10,
  fig.height = 4,
  comment = "#>",
  collapse = TRUE,
  warning = FALSE
)
```

## Acquire the zika GitHub repo

```{r}
library(git2r)
# clone the zika repo if we don't already have it
zikaRepo <- if (!dir.exists("zika")) {
  clone("https://github.com/cdcepi/zika.git", local_path = "zika")
} else {
  repository("zika")
}
zikaRepo
```

## Read in the most recent data

```{r}
library(readr)
dataFiles <- Sys.glob("zika/*/*/data/*.csv")
countries <- unlist(lapply(strsplit(dataFiles, "/"), "[[", 1))
dats <- lapply(dataFiles, function(x) read_csv(x))
```

## Fix some inconsistencies

```{r}
library(dplyr)
library(tidyr)
# Panama dates are reported as M/D/Y
idx <- grep("^Panama", dataFiles)
dats[idx] <- lapply(dats[idx], function(x) {
  x$report_date <- lubridate::mdy(x$report_date)
  x
})

# someone put values as "x*y", lol
is_character <- function(x) typeof(x$value) == "character"
badValues <- which(vapply(dats, is_character, logical(1)))
dats <- lapply(dats, function(x) {
  x$value  <- as.character(x$value)
  x
})

# fix some inconsistent naming (these columns are missing anyway)
nms <- names(dats[[252]])
nms <- sub("time_period_year", "time_period", nms, fixed = TRUE)
nms <- sub("time_period_week", "time_period_type", nms, fixed = TRUE)
names(dats[[252]]) <- nms

# put everything in one big data frame
zika <- bind_rows(dats)

# fix the "unevaluated" multiplication
idx <- grep("\\*", zika$value)
zika$value[idx] <- lapply(
  strsplit(zika$value[idx], "\\*"),
  function(x) prod(as.numeric(x))
)
zika$value <- as.numeric(zika$value)
```

## Janitorial work

```{r}
# hmmm, these variables seem useless, remove them
summary(is.na(zika$time_period))
summary(is.na(zika$time_period_type))
zika <- zika[!grepl("time_period*", names(zika))]

# all municipailities units come from Brazil -- save them in a separate data file
municipalities <- filter(zika, unit == "municipalities")
zika <- filter(zika, unit == "cases")
table(municipalities$country)
devtools::use_data(municipalities, overwrite = TRUE)

# no longer need units (all units are now "cases")
zika$unit <- NULL

# location seems to be in "country-locationA-locationB" format
zika <- zika %>%
  tidyr::separate(
    location,
    into = c('country', 'locationA', 'locationB'),
    sep = '-', fill = 'right', remove = FALSE
  )
```

The script in "code/Plot_cases_by_country/Plot_all.R" uses this table to essentially perform a semi_join on the data, filtering out weird data_field codes and redundant location types.

```{r}
meta <- read_csv("zika/code/Plot_cases_by_country/country_variable_selection.csv")
names(meta) <- sub("country_name", "country", names(meta))
names(meta) <- sub("location_filter", "location_type", names(meta))
meta <- meta %>% 
  transform(confirmed_codes = strsplit(confirmed_codes, ",")) %>%
  unnest() %>%
  transform(suspected_codes = strsplit(suspected_codes, ",")) %>%
  unnest() %>%
  transform(location_type = strsplit(location_type, ",")) %>%
  unnest() %>%
  gather(report_type, data_field_code, confirmed_codes, suspected_codes) %>%
  mutate(report_type = sub("_codes", "", report_type)) %>%
  unique()

vars <- c("country", "data_field_code", "location_type")


zika <- zika %>% 
  semi_join(meta, by = vars) %>% 
  left_join(meta, by = vars) %>% 
  select(-file) %>%
  group_by(location, country, report_date, report_type) %>%
  summarise(value = sum(value)) %>%
  arrange(location, report_type, report_date) %>%
  ungroup()

devtools::use_data(zika, overwrite = TRUE)
```

## Geo-locate locations

Note this takes a while to run, and should be done by hand (for now, at least)

```{r, eval = FALSE}
load_all()
data(zika)
locations <- unique(zika[c("location", "locationA", "locationB", "country")])
#tryCatch(load("../data/latLonDat.rda"), error = function(e) invisible())
#newLocations <- setdiff(locations, unique(latLonDat$location))
#if (length(newLocations)) {
urls <- sprintf(
  "http://maps.googleapis.com/maps/api/geocode/json?address=%s&components=country:%s",
  gsub("_", "+", with(locations, paste(locationA, locationB, sep = "+"))), 
  gsub("_", "+", locations$country)
)
json <- lapply(urls, function(x) jsonlite::fromJSON(x)$results)

latLons <- Map(function(x, y) {
  # location can be a multiple row data frame!
  latlon <- x[["geometry"]][["location"]][1, ]
  latlon$status <- x[["status"]]
  latlon$location <- y
  latlon
}, json, locations$location)

latLonDat <- unique(dplyr::bind_rows(latLons))
#}

# there are some bad locations...
latLonDat %>% filter(is.na(lat))
#>   lat lng                  location
#> 1  NA  NA                     Norte
#> 2  NA  NA                  Nordeste
#> 3  NA  NA                   Sudeste
#> 4  NA  NA                       Sul
#> 5  NA  NA              Centro-Oeste
#> 6  NA  NA Ecuador-Manabi-Rocafuerte
#> 7  NA  NA         Panama_Importados

latLonDat <- latLonDat %>% filter(!is.na(lat))
devtools::use_data(latLonDat, overwrite = TRUE)
```


## Attempt to make sense of field codes

Note: most of this code was done before I discovered the work under code

```{r, eval = FALSE}
# most location types are at the municipality level...
table(zika$location_type)
# because...Colombia!!
table(zika$location_type, zika$country)

# try to infer the disease type reported
is_type <- function(x, type = "zika") {
  as.numeric(grepl(type, x))
}

zika <- zika %>%
  mutate_(
    # remove redundant "zika" references in data field
    #data_field = ~sub("[_]?[Z-z]ika[_]?", "", data_field),
    # replace columbian codes with something meaningful
    data_field = ~recode("CO0001" = "confirmed_laboratory", data_field),
    data_field = ~recode("CO0002" = "confirmed_clinic", data_field),
    data_field = ~recode("CO0003" = "suspected", data_field),
    data_field = ~recode("CO0004" = "suspected_clinic", data_field),
    confirmed = ~ifelse(grepl("confirmed", data_field, ignore.case = T), "confirmed", "not confirmed")
    #gbs = ~is_type(data_field, "gbs"),
    #flavi = ~is_type(data_field, "flavi"),
    #arbovirus = ~is_type(data_field, "arbovirus"),
    #microcephaly = ~is_type(data_field, "microcephaly"),
    #zika = ~is_type(data_field, "zika")
  )

devtools::use_data(zika, overwrite = TRUE)
```

## Filter to the most granular location type within country

Should we automatically show most granular location type when conditioned on a country? (update: for now we're going to use the location types specified in the "zika/code/Plot_cases_by_country/country_variable_selection.csv" file)

```{r, eval = FALSE}
top_locations <- zika %>% 
  group_by(country, location_type) %>% 
  summarise(n = max(length(unique(locationA)), length(unique(locationB)))) %>% 
  slice(which.max(n))
semi_join(zika, top_locations)
```


## Data field guides

```{r, eval = FALSE}
guideFiles <- Sys.glob("zika/*/*[G-g]uide*")
countries <- dirname(guideFiles)
guides <- lapply(guideFiles, function(x) read_csv(x))
guides <- Map(function(x, y) { x$country <- y; x}, guides, countries)
dataGuide <- bind_rows(guides)


# many countries have redundant counts reported on different levels
# for example,
zika %>%
  filter(country == "Brazil") %>%
  group_by(location_type, data_field) %>%
  summarise(n = sum(value))

#>       location_type                        data_field       n
#>          <chr>                                   <chr>   <dbl>
#> 1       country                          zika_reported 1091664
#> 2        region                          zika_reported 1091664
#> 3         state                 microcephaly_confirmed      NA
#> 4         state           microcephaly_fatal_confirmed      NA
#> 5         state                 microcephaly_fatal_not      NA
#> 6         state microcephaly_fatal_under_investigation      NA
#> 7         state                       microcephaly_not      NA
#> 8         state       microcephaly_under_investigation      NA
#> 9         state                          zika_reported 1091664

zika %>%
  filter(country == "United_States") %>%
  group_by(location_type, data_field) %>%
  summarise(n = sum(value))


# We'll focus just on the country level for now


# filter
dataGuide %>%
  filter(country == "Ecuador" &
           data_field %in% c("total_zika_confirmed_cumulative",
                             "total_zika_confirmed_pregnant",
                             ""
           ))

```
